---
title: "QuantAQ Input"
output: html_notebook
---

This is the script for the QuantAQ sensor installed on the
Exterior of Andy Murphy Neighborhood Center. 

Load in the packages that we're using:
```{r}
library(tidyverse)
library(lubridate)
library(openair)
library(googledrive)
```

Read in the csv file. This file has already been downloaded and put into the appropriate folder with the correct name:
```{r}
month_name <- "April 2021" #edit the correct month here

name_file <- "/Mod-PM-00044_Apr2021_AMNC_final.csv" #This is what you should edit each month to change the name of the file
name_file_raw <-  "/Mod-PM-00044_Apr2021_AMNC_raw.csv" #This is what you should edit to change the name of the raw data file

#Concatenate filename to googledrive directory
name_path <- paste("~/Shared with me/Community Science Lab/Monitoring Programs/Continuous Monitoring Sites/Andy Murphy Building/QAQ MOD-PM/Raw Data/Final/", month_name, name_file, sep = "")

#Concatenate filename to googledrive directory - raw data
name_path_raw <- paste("~/Shared with me/Community Science Lab/Monitoring Programs/Continuous Monitoring Sites/Andy Murphy Building/QAQ MOD-PM/Raw Data/Raw/", month_name, name_file_raw, sep = "")

drive_download(file = name_path_raw,
  overwrite = TRUE
)
drive_download(name_path,
  overwrite = TRUE
)


quantaq <- read_csv(name_file) 
quantaq_raw <- read_csv(name_file_raw)
```


Before we drop the device_state column, we can just verify that there are no entries with another device_state besides "ACTIVE". We're expecting that if we are taking a reading that the device is "ACTIVE", but this is just checking in case one time something is different!:
```{r}
q_state <- filter(quantaq,
            device_state != "ACTIVE"
)

q_state

# Do the same for raw data
q_state_raw <- filter(quantaq_raw,
            device_state != "ACTIVE"
)

q_state_raw

```

Drop columns that we are not storing in our cleaned up file. Then rename columns according to lab conventions. Then reorder the dataframe to be in chronological order. Finally, round the seconds of timestamp down to zero for ease in merging this data with other dataframes:
```{r}
quantaq <- quantaq %>% 
  select(-c(X1,device_state,timestamp,pm1_model_id, pm25_model_id, pm10_model_id)) %>% 
  rename( # Rename columns according to lab conventions
    time = timestamp_local 
    # Others to rename? 
  )
  
  quantaq <- quantaq[order(quantaq$time),]
  
  #zero out the seconds of timestamp
  second(quantaq$time) <- 0

  
  # Do the same for raw_data
  quantaq_raw <- quantaq_raw %>% 
  select(-c(X1,device_state,timestamp)) %>% 
  rename( # Rename columns according to lab conventions
    time = timestamp_local,
    raw_table_id = id
    )
  
  quantaq_raw <- quantaq_raw[order(quantaq_raw$time),]
  
  #zero out the seconds of timestamp
  second(quantaq_raw$time) <- 0
  
```

Give yourself a quick look at the data before any joining or cleaning happens:
```{r}
#summary plot
quantaq %>%
  select(
    -c(id,sn,raw_table_id,lat,lon)
  ) %>%
  rename(
    date = time
  ) %>%
summaryPlot(period = "months")
```
And a quick look at some summary statistics for the different variables before any cleaning, scanning to see if the ranges seem in the realm of the possible:
```{r}
summary(quantaq)
```

Isolate the top and bottom .05% in order to scan values to confirm that even the extreme values don't have measurement errors causing issues:
```{r}
# First pm2.5
lower_bound <- quantile(quantaq$pm25, 0.0005)
upper_bound <- quantile(quantaq$pm25, 0.9995)

outlier_ind <- which(quantaq$pm25 < lower_bound | quantaq$pm25 > upper_bound)

quantaq[outlier_ind, ]
```
```{r}
# Then PM 10
lower_bound <- quantile(quantaq$pm10, 0.0005)
upper_bound <- quantile(quantaq$pm10, 0.9995)

outlier_ind <- which(quantaq$pm10 < lower_bound | quantaq$pm25 > upper_bound)

quantaq[outlier_ind, ]
```
Now a quick look at the raw_data before we proceed. There are so many variables in this data, that we want to mostly be verifying that the data is covering the same time period. I'll choose just a handful of variables to display:

```{r}
#summary plot
quantaq_raw %>%
  select(
    time,sample_temp,sample_rh,bin0,bin2,bin3,opcn3_pm25,opcn3_pm10,pm25_env,opcn3_pm10
  ) %>%
  rename(
    date = time
  ) %>%
summaryPlot(period = "months")
```
Join the two sets:

```{r}
quantaq_full <- full_join(quantaq,quantaq_raw, by = "raw_table_id",suffix = c(".final", ".raw"))
```



Clean the variables by dropping rows that have values outside of measurement range (another alternative would be to preserve row but change value to NA):

```{r}
#dropping duplicate columns and a little renaming
quantaq_full <- quantaq_full %>% 
  select(-c(sample_rh.raw,sample_temp.raw,sample_pres.raw,lat.raw,lon.raw, sn.raw)) %>%
  rename(
    time = time.final,
    sn = sn.final,
    sample_temp = sample_temp.final,
    sample_rh = sample_rh.final,
    sample_pres = sample_pres.final
  )


# Marco: need rules confirmed
quantaqc <- quantaq_full %>% 
  filter(
     sample_temp >= -273.15 & sample_temp <= 200,
     sample_rh >= 0 & sample_rh <= 100,
     sample_pres >= 0 & sample_pres <= 2000,
     pm1 >= 0 & pm1 <= 9999,
     pm25 >= 0 & pm25 <= 9999,
     pm10 >= 0 & pm10 <= 9999,
     bin0 >= 0,
     bin1 >= 0,
     bin2 >= 0,
     bin3 >= 0,
     bin4 >= 0,
     bin5 >= 0,
     bin6 >= 0,
     bin7 >= 0,
     bin8 >= 0,
     bin9 >= 0,
     bin10 >= 0,
     bin11 >= 0,
     bin12 >= 0,
     bin13 >= 0,
     bin14 >= 0,
     bin15 >= 0,
     bin16 >= 0,
     bin17 >= 0,
     bin18 >= 0,
     bin19 >= 0,
     bin20 >= 0,
     bin21 >= 0,
     bin22 >= 0,
     bin23 >= 0,
     opcn3_temp >= -273.15 & opcn3_temp <= 200,
     opcn3_rh >= 0 & opcn3_rh <= 100,
     pm1_env >= 0 & pm1_env <= 9999,
     pm25_env >= 0 & pm25_env <= 9999,
     pm10_env >= 0 & pm10_env <= 9999,
     neph_bin0 >= 0,
     neph_bin1 >= 0,
     neph_bin2 >= 0,
     neph_bin3 >= 0,
     neph_bin4 >= 0,
     neph_bin5 >= 0,
     flag == 0
  )

```

```{r}
glimpse(quantaqc)
```

Number of rows dropped and % of rows dropped by cleaning:
```{r}
paste("Number of rows dropped:", nrow(quantaq_full) - nrow(quantaqc))
p <- (nrow(quantaq_full) - nrow(quantaqc))/nrow(quantaq_full)
paste("% of rows dropped:", p)
```



And finally export the clean csv file to the designated storage folder. Right now I have it exporting to a csv file. You could consider writing to a feather file format as well:


```{r}
# Take the orginal file name and insert "clean" label in file name
path_name <- str_sub(name_file,1,nchar(name_file)-4) %>%
  paste("_clean.csv", sep = "")

write_csv(quantaqc, file = path_name)


#Concatenate file name to GoogleDrive path location
googlepath <- paste( "~/Community Science Lab/Monitoring Programs/Continuous Monitoring Sites/Andy Murphy Building/QAQ MOD-PM/Processed Data", path_name, sep = "")

#uploading processed dataframe to google drive
processedDF <- drive_upload(path_name, type = "csv", 
                  path = googlepath, overwrite = TRUE)
```


Beta version of interacting with the google drive. Let's discuss before this gets used:

```{r}
#download master file
master <- drive_download(
  "~/Community Science Lab/Monitoring Programs/Continuous Monitoring Sites/Andy Murphy Building/QAQ MOD-PM/Master/AMNC_quantaq_master.csv",
  overwrite = TRUE
)

#binding the new month's data to master and creating new master csv file
fm <- read_csv(master)
newmaster <- bind_rows(fm, quantaqc)
write_csv(newmaster, file ="AMNC_quantaq_master.csv") #verify that it overwrites

#uploading new updated master file to google drive
nmaster <- drive_upload("AMNC_quantaq_master.csv", type = "csv", 
                  path = "~/Community Science Lab/Monitoring Programs/Continuous Monitoring Sites/Andy Murphy Building/QAQ MOD-PM/Master/AMNC_quantaq_master.csv", overwrite = TRUE)


```

